The Effect of a Second AIDS Defining Illness on Survival in the MACS

### 2015-05-09 paper.Rmd
Many things...

---
title:  "The Effect of a Second AIDS Defining Illness on Survival in the MACS"
author: 
- "[Jacob Carey](mailto:jcare15@jhu.edu)"
- "[Matt Murrill](mmurril1@jhmi.edu)"
- "[Kim Kunwha](khkim@jhu.edu)"
date:   "`r Sys.Date()`"
output: pdf_document
abstract: |
  We seek to investigate the effect of a second AIDS defining illness on survival. We included people who are seroprevalent as well as seroconverters. We addressed confounding using propensity scores.
---

@@

```{r libs}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(survival)
theme_set(theme_classic())
```

We read in the data.
```{r data}
# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv", 
                       progress=FALSE) %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           age=as.Date("1960-01-01") + DOB,
           exit=ifelse(!died, as.Date("1994-12-31"), death.date),
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           censor.date=as.Date("1970-01-01")+censor.date,
           dead.at.censor=(death.date < as.Date("1994-12-31"))) %>%
    select(STUDYID, 
           aids.date,
           sec.aids.date,
           frst.aids.diag, 
           censor.date, 
           sec.aids.diag,
           dead.at.censor, 
           death.date) %>%
    filter(year(aids.date) < 2100)

# read in visit level data
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # keep MACS only
    filter(COHORT==1) %>%
    select(STUDYID) %>%
    distinct(STUDYID)

covariates <- covariates %>%
    inner_join(visits)
```

```{r fig1}
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # put visits in correct 1,2, 3,... form
    mutate(visit=VISIT / 10,
           visit.date=mdy(VDATE),
           haart.era.visit=year(visit.date) <= 1994) %>%

    # join to covariates
    left_join(covariates, by=c("STUDYID")) %>%

    # keep MACS only
    filter(COHORT==1,
           # keep people with first ADI only
           frst.aids.diag != 0,
           # restrict to haart era visits right now
           haart.era.visit == TRUE) %>%
    select(id=STUDYID, center=CENTER, visit:dead.at.censor)

# only for pre haart!
# need to do again for all data
data.plot <- visits %>%
    group_by(id) %>%
    summarise(sec.aids=min(sec.aids.diag)) %>%
    mutate(sec.aids=plyr::revalue(as.character(sec.aids),
                                  c("0"="No Aids",
                                    "1"="PCP",
                                    "2"="KS",
                                    "3"="NHL",
                                    "4"="Neuro",
                                    "5"="Bacteria OI",
                                    "6"="Fungal OI",
                                    "7"="Protozoa",
                                    "8"="CMV/HSV",
                                    "9"="Wasting",
                                    "10"="Recurrent PNA",
                                    "11"="Cervical Ca"),
                                  warn_missing=FALSE))
ggplot(data.plot, aes(x=sec.aids)) + 
    geom_histogram() +
    xlab("Type of Secondary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Figure 1") +
    scale_x_discrete(limits=names(sort(-table(data.plot$sec.aids)))) +
    theme(axis.text.x = element_text(angle=-90))
```

We fit a Cox proportional hazards model to the pre haart era data. We allowed secondary aids diagnosis to be time varying. 
```{r naive-survival}
# http://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf
# https://stat.ethz.ch/R-manual/R-devel/library/survival/html/coxph.html
data <- covariates %>%
    filter(year(aids.date) <= 1994) %>%
    rowwise() %>%
    transmute(start=aids.date,
              stop=min(censor.date, as.Date("1994-12-31")),
              sec.aids=sec.aids.date,
              status=(death.date <= as.Date("1994-12-31"))) %>%
    mutate_each(funs(as.numeric), start, stop, sec.aids) %>%
    # there are a few obs coming up where secondary and primary are 
    # diagnosed at the same time.
    # let's remove those.
    filter(stop > start,
           !is.na(sec.aids))

pre.sec <- data %>%
    rowwise() %>%
    mutate(stop=min(stop, sec.aids),
           sec.aids=0) %>%
    # remove those with same date diag of first and sec aids
    filter(start < stop)

knitr::kable(coef(summary(naive.model)),
             caption="Naive Model (All Data)")
```

```{r all-data}
# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           dob=as.Date("1960-01-01") + DOB,
           age.at.aids=difftime(aids.date, dob, unit="weeks") / 52.25,
           age.at.aids=as.numeric(age.at.aids),
           exit=death.date,
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           censor.date=as.Date("1970-01-01")+censor.date) %>%
    select(STUDYID, 
           aids.date,
           sec.aids.date,
           frst.aids.diag, 
           censor.date, 
           sec.aids.diag,
           death.date,
           race=RACE,
           age.at.aids) %>%
    filter(year(aids.date) < 2100) %>%
    mutate(haart.era=(year(aids.date) > 1993),
           aids.year=year(aids.date),
           aids.month=month(aids.date),
           race=ifelse(race == 1, "white", "non-white"))

# what are the important confoudners that we can include for the PH model?
# what about propensity score?

# what about therapy?
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    # keep MACS only
    filter(COHORT == 1) %>% 
    group_by(STUDYID) %>%
    summarise(drug.use=sum(POTUSE, CRACKUSE, COCUSE, HERUSE, AMPHUSE, 
                           POPUSE, na.rm=TRUE) > 1,
              depression=mean(CESD, na.rm=TRUE) < ifelse(drug.use, 
                                                         16, 23)) %>%
    ungroup() %>%
    # remove the 8 missing depression status
    filter(!is.na(depression))

@@

# add data together
covariates <- covariates %>%
    inner_join(visits) 
```

We also created a naive model that includes **all** MACS data,
but does not adjust for any confounders, or the pre / post HAART era.

```{r all-naive-model}
data <- covariates %>%
    rowwise() %>%
    mutate(start=aids.date,
           stop=censor.date,
           sec.aids=sec.aids.date,
           status=(death.date < as.Date("2100-01-01"))) %>%
    mutate_each(funs(as.numeric), start, stop, sec.aids) %>%
    # there are a few obs coming up where secondary and primary are 
    # diagnosed at the same time.
    # let's remove those.
    filter(stop > start,
           !is.na(sec.aids))

pre.sec <- data %>%
    rowwise() %>%
    mutate(stop=min(stop, sec.aids),
           sec.aids=0) %>%
    # remove those with same date diag of first and sec aids
    filter(start < stop)

@@

@@

```{r all-adjusted}
adjusted.model <- coxph(Surv(start, stop, status) ~ sec.aids +
                        factor(race) + age.at.aids + drug.use +
                        aids.year + depression + cd4,
                        data=time.varying)

@@

Finally, we use propensity score matching to account for confounding bias.

```{r propensity-score}
caliper <- 0.07
set.seed(90210)
match.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                              age.at.aids + drug.use +
                              aids.year + depression + cd4,
                              data=time.varying,
                              distance="logit",
                              method="nearest",
                              caliper=caliper,
                              discard="both")

# ## let's check out the diagnostics

# ## numerical check
match.check <- summary(match.set, standardize=TRUE)

# ## graphical check
plot(match.set, type="hist")
# plot(match.set, type="jitter", interactive=FALSE)

matched.dat <- MatchIt::match.data(match.set,"all")

ph.match <- coxph(Surv(start, stop, status) ~ sec.aids +
                  factor(race) + age.at.aids + drug.use +
                  aids.year + depression + cd4,
   				  data=matched.dat,
                  method="efron")
```

There were `r length(unique(matched.dat$STUDYID))` individuals in the 
matched data set of the orginal `r nrow(data)`. There are 
`r sum(matched.dat$sec.aids)` events in the matched data set.
We perform a check on the data and see the following summaries. We used a
caliper of `r caliper`.

```{r}
knitr::kable(match.check$sum.all[,1:4],
             caption="Summary of balance for all data")
knitr::kable(match.check$sum.matched[,1:4],
             caption="Summary of balance for matched data")
knitr::kable(match.check$reduction,
             caption="Percent Balance Improvement")
knitr::kable(match.check$nn,
             caption="Sample Sizes")
```

In recalculating our proportional hazards model using propensity score
in order to handle confounding bias, we see the following results.

```{r}
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
caliper <- 0.5
```

Next, we performed the same PH model using propensity scores, but with a higher
caliper (`r caliper`).

```{r propensity-score-2}
set.seed(90210)
match.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                              age.at.aids + drug.use +
                              aids.year + depression + cd4,
                              data=time.varying,
                              distance="logit",
                              method="nearest",
                              caliper=caliper,
                              discard="both")

# ## let's check out the diagnostics

# ## numerical check
match.check <- summary(match.set, standardize=TRUE)

# ## graphical check
plot(match.set, type="hist")
# plot(match.set, type="jitter", interactive=FALSE)

matched.dat <- MatchIt::match.data(match.set,"all")

ph.match <- coxph(Surv(start, stop, status) ~ sec.aids +
                  factor(race) + age.at.aids + drug.use +
                  aids.year + depression + cd4,
   				  data=matched.dat,
                  method="efron")
```

There were `r length(unique(matched.dat$STUDYID))` individuals in the 
new matched data set of the orginal `r nrow(data)`. There are 
`r sum(matched.dat$sec.aids)` events in the matched data set.
We perform a check on the data and see the following summaries.

```{r}
knitr::kable(match.check$sum.all[,1:4],
             caption="Summary of balance for all data")
knitr::kable(match.check$sum.matched[,1:4],
             caption="Summary of balance for matched data")
knitr::kable(match.check$reduction,
             caption="Percent Balance Improvement")
knitr::kable(match.check$nn,
             caption="Sample Sizes")
```

In our second PH model with propensity scores, we see the following results.

```{r}
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
caliper <- 0.5
```

Finally, we used a subclassification method for our propensity score.

```{r subclass}
sub.set <- MatchIt::matchit(sec.aids ~ factor(race) + 
                            age.at.aids + drug.use +
                            aids.year + depression + cd4,
                            data=time.varying,
                            method="subclass",
                            sub.by="all",
                            subclass=5)

# summary(sub.set,standardize=T)
plot(sub.set,type="jitter",interactive=F)
# plot(summary(sub.set,standardize=T),interactive=F)

sub.dat <- MatchIt::match.data(sub.set, "all")
ph.match <- coxph(Surv(start, stop, status) ~ sec.aids + strata(subclass),
   				  data=sub.dat,
                  method="efron")
knitr::kable(coef(summary(ph.match)),
             caption="PH Model (Propensity Scores)")

knitr::kable(cox.zph(ph.match)$table,
             caption="PH Assumption Test")
```

```{r graphs}
plot(survfit(ph.match), main="Final PH Model Using Subclassification")
```

>>>>

---
title:  "The Effect of a Second AIDS Defining Illness on Survival in the MACS"
author: 
- "[Jacob Carey](mailto:jcare15@jhu.edu)"
- "[Matt Murrill](mailto:mmurril1@jhmi.edu)"
- "[Kim Kunwha](mailto:khkim@jhu.edu)"
date:   "`r Sys.Date()`"
output: word_document
abstract: |
  We seek to investigate the effect of a second AIDS defining illness on survival. We included people who are seroprevalent as well as seroconverters. We addressed confounding using propensity scores.
---

@@

```{r libs}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(survival)
theme_set(theme_classic())
```

```{r all-data}
# bins for 2nd ADI
# PCP, KS, CMV, NHL, neuro, other
map <- c("0"="No Aids",
         "1"="PCP",
         "2"="KS",
         "3"="NHL",
         "4"="Neuro",
         "5"="Other",
         "6"="Other",
         "7"="Other",
         "8"="CMV/HSV",
         "9"="Other",
         "10"="Other",
         "11"="Other")

map.2 <- c("1"="Baltimore",
           "2"="Chicago",
           "3"="Pittsburgh",
           "4"="Los Angeles")

# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # aids interval
           aids.interval=ifelse(year(aids.date) < 1990, "[84, 90)", 
                         ifelse(year(aids.date) < 1993, "[90, 93)",
                         ifelse(aids.date < as.Date("1995-07-01"), 
                                "[93, 07-95)",
                                "[07-95, 96)"))),
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # sec aids interval
           seca.interval=ifelse(year(sec.aids.date) < 1990, "[84, 90)", 
                         ifelse(year(sec.aids.date) < 1993, "[90, 93)",
                         ifelse(sec.aids.date < as.Date("1995-07-01"), 
                                "[93, 07-95)",
                                "[07-95, 96)"))),
           # what was the first aids diagnosis
           frst.aids.diag=plyr::revalue(as.character(AIDSDIAG), map, 
                                        warn_missing=FALSE),
           # what was the first aids diagnosis
           sec.aids.diag=plyr::revalue(as.character(SECAIDSDIAG), map, 
                                       warn_missing=FALSE),
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 1996,
           dob=as.Date("1960-01-01") + DOB,
           age.at.aids=difftime(aids.date, dob, unit="weeks") / 52.25,
           age.at.aids=as.numeric(age.at.aids)) %>%
    select(STUDYID, 
           aids.date,
           aids.interval,
           sec.aids.date,
           seca.interval,
           had.sec,
           frst.aids.diag, 
           sec.aids.diag,
           death.date,
           race=RACE,
           age.at.aids) %>%
    # only keep pts who developed aids before 1996
    filter(year(aids.date) < 1996) %>%
    mutate(aids.year=year(aids.date),
           aids.month=month(aids.date),
           race=ifelse(race == 1, "white", "non-white")) %>%
    rowwise() %>%
    mutate(sec.aids.date=min(sec.aids.date, 
                             as.Date("1995-12-31"))) %>%
    ungroup() %>%
    mutate(sec.aids.date=ifelse(had.sec, sec.aids.date, 
                                as.Date("2100-01-01")),
           sec.aids.date=as.Date(sec.aids.date, "1970-01-01"),
           death.date=ifelse(year(death.date) >= 1996,
                             as.Date("2100-01-01"),
                             death.date),
           death.date=as.Date(death.date, "1970-01-01")) %>%
    rowwise() %>%
    mutate(censor.date=min(death.date, as.Date("1995-12-31")),
           censor.date=as.Date(censor.date, "1970-01-01")) %>%
    ungroup() %>%
    tbl_df()

# get CENTER
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>%
    filter(COHORT == 1) %>%
    count(STUDYID, CENTER) %>%
    group_by(STUDYID) %>%
    # just consider the most visited center
    filter(n==max(n)) %>%
    # there are a few ties so let's just pick the first one
    distinct(STUDYID) %>%
    # forget n
    select(-n) %>%
    mutate(CENTER=plyr::revalue(as.character(CENTER), 
                                map.2, warn_missing=FALSE))

# get drug use and depression status
visits <- read_csv("data/hivvert-data04.csv", progress=FALSE,
                    col_types="iiciiididdiiiiiiiddiiiiiiiiiiiidc") %>% 
    # keep MACS only
    filter(COHORT == 1) %>% 
    group_by(STUDYID) %>%
    summarise(drug.use=sum(CRACKUSE, COCUSE, HERUSE, 
                           AMPHUSE, na.rm=TRUE) > 1,
              depression=mean(CESD, na.rm=TRUE) > 16) %>%
    ungroup() %>%
    # remove the 8 missing depression status
    filter(!is.na(depression)) %>%
    # join in center
    left_join(visits)

@@

# add data together
covariates <- covariates %>%
    inner_join(visits) %>%
    filter(!is.na(cd4))
```

```{r graphs-explr}
data.plot <- covariates %>%
    mutate(adi=plyr::revalue(as.character(frst.aids.diag),
                             map, warn_missing=FALSE),
           sec.adi=plyr::revalue(as.character(sec.aids.diag),
                                 map, warn_missing=FALSE)) %>%
    select(adi,
           sec.adi,
           aids.date,
           sec.aids.date,
           death.date)

ggplot(data.plot, aes(x=adi)) + 
    geom_histogram() +
    xlab("Type of Primary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Distibution of Primary ADI's") +
    scale_x_discrete(limits=names(sort(-table(data.plot$adi)))) +
    theme(axis.text.x = element_text(angle=-90))
```

\newpage

red line is mean.

```{r graphs-2}
data.plot.1 <- data.plot %>%
    filter(year(death.date) < 2100) %>%
    mutate(time.to.death=difftime(death.date, aids.date, units="weeks"),
           time.to.death=as.numeric(time.to.death) / 52.25)

vline <- data.plot.1 %>%
    group_by(adi) %>%
    summarise(mean.time=mean(time.to.death),
              count=n())

ggplot(data.plot.1, aes(x=time.to.death)) +
    geom_density() +
    geom_vline(data=vline,
               aes(xintercept=mean.time),
               colour="red",
               linetype="dashed") +
    xlab("Years from AIDS Diagnosis to Death") +
    ggtitle("Distribution of Deadliness of ADI") +
    facet_wrap(~adi)

data.plot.2 <- data.plot %>%
    filter(year(death.date) < 2100,
           year(sec.aids.date) < 2100) %>%
    mutate(time.to.death=difftime(death.date, sec.aids.date, units="weeks"),
           time.to.death=as.numeric(time.to.death) / 52.25)

vline <- data.plot.2 %>%
    group_by(sec.adi) %>%
    summarise(mean.time=mean(time.to.death),
              count=n())

ggplot(data.plot.2, aes(x=time.to.death)) +
    geom_density() +
    geom_vline(data=vline,
               aes(xintercept=mean.time),
               colour="red",
               linetype="dashed") +
    xlab("Years from Second AIDS Diagnosis to Death") +
    ggtitle("Distribution of Deadliness of Secondary ADI") +
    facet_wrap(~sec.adi)
```

\newpage

We created a naive model that includes incident and prevalent cases
diagnoses before 1996 and administratively censored them at 1996.
We did not adjust for any confounders.

```{r all-naive-model}
data <- covariates %>%
    rowwise() %>%
    mutate(start=aids.date,
           stop=censor.date,
           sec.aids=sec.aids.date,
           status=(death.date < as.Date("2100-01-01"))) %>%
    mutate_each(funs(as.numeric), start, stop, sec.aids) %>%
    # remove those with NA for sec aids date
    # and those with aids diag and death on same day
    filter(!is.na(sec.aids),
           aids.date != death.date)

pre.sec <- data %>%
    rowwise() %>%
    mutate(stop=min(stop, sec.aids),
           sec.aids=0) %>%
    # remove those with sec aids == start aids
    # well put them in post sec
    filter(stop > start)

@@

\newpage

@@

```{r all-adjusted}
adjusted.model <- coxph(Surv(start, stop, status) ~ sec.aids +
                        factor(race) + age.at.aids + drug.use +
                        depression + cd4 + 
                        factor(CENTER) + aids.interval,
                        data=time.varying)

@@

\newpage

Finally, we use an IPW approach to adjust for confounding bias.

```{r ipw}
cpoe.model <- glm(sec.aids ~ factor(race) + age.at.aids +
                  drug.use + depression + cd4 + 
                  factor(CENTER) + aids.interval, 
                  data=time.varying,
                  family=binomial)

mpoe.model <- glm(sec.aids ~ 1, data=time.varying, family=binomial)

time.varying <- time.varying %>%
    mutate(sec.aids=as.integer(sec.aids),
           cpoe=fitted(cpoe.model),
           mpoe=fitted(mpoe.model),
           w=ifelse(sec.aids == 1, 1 / cpoe, 1 / (1 - cpoe)),
           sw=ifelse(sec.aids == 1, mpoe / cpoe, (1 - mpoe) / (1 - cpoe)))

data.graph <- time.varying %>%
    select(w:sw) %>%
    gather(weight, value, w:sw)

ggplot(data.graph, aes(x=value)) +
    geom_density() +
    xlab("Weight") +
    ylab("") +
    ggtitle("Distribution of weights") +
    facet_wrap(~weight)

ggplot(time.varying, aes(x=factor(sec.aids), y=log(sw))) +
    geom_boxplot() +
    xlab("Secondary AIDS Event") +
    ylab("Log(SW)") +
    ggtitle("Spread of Log SW by Secondary AIDS Event")

ph.sw <- coxph(Surv(start, stop, status) ~ sec.aids + cluster(STUDYID),
               data=time.varying,
               weights=sw ,
               method="efron")

knitr::kable(coef(summary(ph.sw)),
             caption="PH Model Weighted by SW")

knitr::kable(cox.zph(ph.sw)$table,
             caption="PH Assumption Test")
```

```{r graphs-final}
# plot(survfit(ph.sw), main="Final PH Model Using IPW")
```

Survival in months.
```{r table1}
tbl.func <- function(data, col) {
    row.data <- data %>%
        group_by_(col) %>%
        summarise(col=col,
                  n=n(),
                  deaths=sum(died),
                  death.pct=mean(died),
                  surv.25=quantile(survival.time, 0.25, na.rm=TRUE),
                  surv.med=quantile(survival.time, 0.5, na.rm=TRUE),
                  surv.75=quantile(survival.time, 0.75, na.rm=TRUE))

    names(row.data)[1] <- "characteristic"

    return(row.data)
}

tbl.pre <- data %>%
    transmute(site=CENTER,
              race=race,
              age.ge.40=ifelse(age.at.aids >= 40,
                               "age >= 40", "age < 40"),
              aids.year=cut(year(aids.date), c(1984, 1988, 1992, 1995), 
                                 include.lowest=TRUE),
              cd4.base=cut(cd4, c(0, 50, 200, Inf), include.lowest=TRUE),
              initial.aids=frst.aids.diag,
              second.aids=sec.aids.diag,
              depressed=ifelse(depression, "depressed", "not depressed"),
              drug.user=ifelse(drug.use, "Drug User", "Not a Drug User"),
              site=CENTER,
              died=year(death.date) < 2100,
              survival.time=difftime(death.date, aids.date, 
                                     units="weeks") / 52.25 * 12,
              survival.time=as.numeric(survival.time),
              survival.time=ifelse(died, survival.time, NA))

all.data <- tbl.pre %>%
    summarise(characteristic="all",
              col="all",
              n=n(),
              deaths=sum(died),
              death.pct=mean(died),
              surv.25=quantile(survival.time, 0.25, na.rm=TRUE),
              surv.med=quantile(survival.time, 0.5, na.rm=TRUE),
              surv.75=quantile(survival.time, 0.75, na.rm=TRUE))

table.1 <- bind_rows(all.data,
                     tbl.func(tbl.pre, "race"),
                     tbl.func(tbl.pre, "age.ge.40"),
                     tbl.func(tbl.pre, "aids.year"),
                     tbl.func(tbl.pre, "cd4.base"),
                     tbl.func(tbl.pre, "initial.aids"),
                     tbl.func(tbl.pre, "second.aids"),
                     tbl.func(tbl.pre, "depressed"),
                     tbl.func(tbl.pre, "drug.user"),
                     tbl.func(tbl.pre, "site"))

knitr::kable(table.1,
             digits=4,
             caption="Table 1")
```

Keri's recommendations:  
    1. ~~Origin being primary diagnosis is fine~~  
    2. ~~person time since prim diag is fine~~  
    3. be careful w/ term time varying exposure  
    4. ~~covarites: race is fine, age is fine, remove marijauan and poppers from drug~~  
    5. ~~depression use 23 period~~  
    6. should depression be mean?  
    7. ~~CD4 is great~~  
    8. ~~year of aids diagnosis? restrict to 1996. Censor study at 1996. could categorize year (<90, 90-93, 93-Jul 95, >Jul 95)~~  
    9. ~~put people back in that had sec aids at baseline.~~  
    10. timing of secondary aids diagnosis? distribution of prim aids to sec. aids infection?  
    11. sens. look at dist of survival by type. pick a cutoff that we think is  meaningful. type of primary. i.e. higher median survival, lower median survival.  
    12. ~~switch propensity scores to ipw~~  
    13. combinations of prim and sec aids impact. could explore with interaction.  
    14. ~~Add site~~
